---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```
## Loading and preprocessing the data
```{r, echo =TRUE}
activity <- read.table("activity.csv", sep =",", header=TRUE)
summary(activity)
meaningful_steps <- activity[!is.na(activity$steps),]
meaningful_steps$date <- as.Date(meaningful_steps$date)
by_day <- group_by(meaningful_steps, date)
steps_by_day <- summarise(by_day, total = sum(steps))
summary(steps_by_day)
```


## What is mean total number of steps taken per day?
Histogram of the steps per day:

```{r, echo=TRUE}
ggplot(steps_by_day, aes(total)) + geom_histogram()
```

Mean of steps taken per day is 10,766 and median is 10,765.


## What is the average daily activity pattern?


###Time series Plot:
```{r, echo = TRUE}
steps_by_interval <- aggregate(steps ~ interval, meaningful_steps, mean)
plot(steps_by_interval$interval, steps_by_interval$steps, type = 'l', main = "Averag number of steps over all th days", xlab = "Interval", ylab = "Average number of steps")
```


## Imputing missing values

Total number of NA:
```{r, echo = TRUE}
sum(is.na(activity$steps))
impute_activity <- activity
impute_activity$steps[is.na(impute_activity$steps)] <- mean(impute_activity$steps, na.rm = T)
sum(is.na(impute_activity$steps))
impute_activity$date <- as.Date(impute_activity$date)
impute_by_day <- group_by(impute_activity, date)
impute_steps_by_day <- summarise(impute_by_day, total = sum(steps))
summary(impute_steps_by_day)
ggplot(impute_steps_by_day, aes(total)) + geom_histogram()
```

Mean=Median after imputing.
Mean is 10,766 steps and median is 10,766 steps.

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo = TRUE}
impute_activity$day <- weekdays(as.Date(impute_activity$date))
impute_activity$weekday[(impute_activity$day == "Saturday" | impute_activity$day == "Sunday")] <- "weekend"
impute_activity$weekday[!(impute_activity$day == "Saturday" | impute_activity$day == "Sunday")] <- "weekdays"
impute_activity$weekday <- as.factor(impute_activity$weekday)
summary(impute_activity)

#time seires plot on 5 minute interval(x-axis) and avg number of steps taken, averaged across all weekdays or weekends.
ggplot(impute_activity, aes(interval, steps)) + geom_smooth(na.rm = TRUE) + facet_wrap(~weekday)
```